Finiteness in Cubical Type Theory

### 2020-01-18 paper.tex
more

\input{preamble}
\title{Finiteness in Cubical Type Theory}
\author{Donnacha Oisín Kidney}
\authorrunning{D. O. Kidney}
\institute{University College Cork \and
  \email{o.kidney@cs.ucc.ie}}
\begin{document}
\maketitle
\begin{abstract}
  We study five different notions of finiteness in Cubical Type Theory and prove
  the relationship between them.
  In particular we show that any totally ordered Kuratowski finite type is
  manifestly Bishop finite.

@@

Homotopy Type Theory \cite{hottbook} adds another wrinkle to the story.
Firstly, in HoTT we cannot assume that every type is a (homotopy) set: this
means that the finiteness predicates above can be further split into versions
which apply to sets only, and those that apply to all types.
Secondly, HoTT gives us a principled and powerful way to construct quotients,
allowing us to regain some of the flexibility of classical definitions by
``forgetting'' the parts of a proof we would be forced to remember in MLTT.
\subsection{Contributions}
\begin{itemize}
  \item Relation between different versions of finiteness, especially card to
    bishop.
  \item Topos for Kuratowski
  \item Countable and W-topos
  \item Search
\end{itemize}
\subsection{Related Work}
\subsection{Notation}

>>>>

\input{preamble}
\title{Finiteness in Cubical Type Theory}
\author{Donnacha Oisín Kidney}
\authorrunning{D. O. Kidney}
\institute{University College Cork \and
  \email{o.kidney@cs.ucc.ie}}
\author{Gregory Provan}
\institute{University College Cork \and
  \email{g.provan@cs.ucc.ie}}
\begin{document}
\maketitle
\begin{abstract}
  We study five different notions of finiteness in Cubical Type Theory and prove
  the relationship between them.
  In particular we show that any totally ordered Kuratowski finite type is
  manifestly Bishop finite.

@@

Homotopy Type Theory \cite{hottbook} adds another wrinkle to the story.
Firstly, in HoTT we cannot assume that every type is a (homotopy) set: this
means that the finiteness predicates above can be further split into versions
which apply to sets only, and those that apply to all types.
Secondly, HoTT gives us a principled and powerful way to construct quotients,
allowing us to regain some of the flexibility of classical definitions by
``forgetting'' the parts of a proof we would be forced to remember in MLTT.

Finally, the other important property of constructive mathematics is that we can
actually \emph{compute} with proofs.
Cubical Type Theory \cite{cohenCubicalTypeTheory2016}, and its implementation in
Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, realise this property
even in the presence of univalence, giving computational content to programs
written in HoTT.
\subsection{Contributions}
In this work we will examine five notions of finiteness in Homotopy Type Theory:
the relationships between them, and their topos-theoretic characterisation.
We also briefly examine a definition for countable sets, comparing it to the
manifestly Bishop finite sets.
Our work is formalised in Cubical Agda, where we also develop a library for
proof search based on the finiteness predicates.

The finiteness predicates we are interested in are organised in
Figure~\ref{finite-classification}.
We will provide a definition for each, and prove each relation to the other
finiteness predicates.
In terms of Figure~\ref{finite-classification}, this means providing a function
to inhabit each arrows.

We will then characterise each predicate as some form of topos.
Our proofs follow the structure of \cite[Chapters 9, 10]{hottbook} and
\cite{rijkeSetsHomotopyType2015}.
This means we will define first the precategory of sets, and then the category
of sets, and prove the required closures and limits to get us to a topos.
We say ``some form of'' topos here because of course the category of sets in
HoTT do \emph{not}, in fact, form a topos, rather a \(\Pi W\)-pretopos.

\begin{itemize}
  \item Relation between different versions of finiteness, especially card to
    bishop.
  \item Topos for Kuratowski
  \item Countable and W-topos
  \item Search
\end{itemize}
\subsection{Related Work}
\subsection{Notation}
